---
title: "Webinar: Propensity Score Analysis in Healthcare Data (Part 3: PS matching, more options)"
author: "Ehsan Karim, ehsan.karim@ubc.ca"
date: "`r format(Sys.time(), '%d %B %Y')`"
always_allow_html: yes
header-includes:
- \usepackage{float}
- \usepackage{booktabs}
- \floatplacement{figure}{H}
- \usepackage{fancyhdr,textcomp,xcolor}
- \pagestyle{fancy}
- \fancyhead[LE,RO]{\texttt{\tiny{\textcolor{gray}{Do not use, share or revise without explicit written permission from the copyright owner} \\ \textcolor{gray}{Ehsan Karim} \textcolor{red}{\textcopyright} \textcolor{gray}{\the\year} \textcolor{blue}{ ehsan.karim@ubc.ca}}}}
output:
  html_document:
    fig_caption: yes
    keep_md: yes
    toc: yes
    number_sections: true
    toc_depth: 2
    toc_float: 
      collapsed: true
      smooth_scroll: true
    theme: lumen
    highlight: textmate
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
  slidy_presentation:
    toc: yes  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(kableExtra)
require(Hmisc)
require(tableone)
require(Publish)
require(jtools)
require(broom.mixed)
require(MatchIt)
require(survey)
```

# Read data

```{r load, cache=TRUE}
analytic.data <- readRDS("data/RHC.Rds")
baselinevars <- c("age","sex", "race","Disease.category", 
                  "Cancer", "DNR.status", "APACHE.III.score",
                  "Pr.2mo.survival","No.of.comorbidity",
                  "ADLs.2wk.prior","DASI.2wk.prior","Temperature",
                  "Heart.rate", "Blood.pressure","Respiratory.rate",
                  "WBC.count","PaO2.by.FIO2","PaCO2","pH",
                  "Creatinine","Albumin","GComa.Score")
```

# Propensity score (PS) analysis

## Table 1 (stratified by exposure)

```{r tab1, cache=TRUE, echo= TRUE,results="hide"}
require(tableone)
tab1 <- CreateTableOne(vars = baselinevars,
               data = analytic.data, strata = "RHC", includeNA = TRUE, 
               test = FALSE, smd = TRUE)
tab1k <- kable(print(tab1, showAllLevels = TRUE, smd = TRUE)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = FALSE)
```

```{r tab1b, cache=TRUE}
tab1k
```

## PS modelling

```{r ps, cache=TRUE}
ps.formula <- as.formula(I(RHC == "RHC") ~ age+sex+race+Disease.category+ 
                                Cancer+DNR.status+APACHE.III.score+
                                Pr.2mo.survival+No.of.comorbidity+
                                ADLs.2wk.prior+DASI.2wk.prior+Temperature+
                                Heart.rate+Blood.pressure+Respiratory.rate+
                                WBC.count+PaO2.by.FIO2+PaCO2+pH+
                                Creatinine+Albumin+GComa.Score)
# fit logistic regression
PS.fit <- glm(ps.formula,family="binomial", data=analytic.data)
# extract propensity scores
analytic.data$PS <- predict(PS.fit, newdata = analytic.data, type="response")
# summarize propensity scores
summary(analytic.data$PS)
# summarize propensity scores by exposure group
tapply(analytic.data$PS, analytic.data$RHC, summary)
# plot propensity scores by exposure group
plot(density(analytic.data$PS), col = "white", type = "n", main = "")
lines(density(analytic.data$PS[analytic.data$RHC == "RHC"]), 
     col = "red")
lines(density(analytic.data$PS[analytic.data$RHC == "No RHC"]), 
      col = "blue", lty = 2)
legend("topright", c("RHC","No RHC"), 
       col = c("red", "blue"), lty=1:2)
```

# PS Matching 

We follow 4 steps as described in Austin 2011.

## Step 1
Specify the propensity score model to estimate propensity scores

```{r ps1, cache=TRUE}
ps.formula <- as.formula(I(RHC == "RHC") ~ age+sex+race+Disease.category+ 
                                Cancer+DNR.status+APACHE.III.score+
                                Pr.2mo.survival+No.of.comorbidity+
                                ADLs.2wk.prior+DASI.2wk.prior+Temperature+
                                Heart.rate+Blood.pressure+Respiratory.rate+
                                WBC.count+PaO2.by.FIO2+PaCO2+pH+
                                Creatinine+Albumin+GComa.Score)
```

## Step 2

```{r ps2, cache=TRUE}
require(MatchIt)
set.seed(123)
# This function fits propensity score model (using logistic 
# regression as above) when specified distance = 'logit'
# performs nearest-neighbor (NN) matching, 
# without replacement 
# with caliper = .2*SD of propensity score  
logitPS <-  -log(1/analytic.data$PS - 1) # calculate the logit of the propensity score
hist(logitPS)
# within which to draw control units 
# with 1:2 ratio (not pair-matching)
match.obj <- matchit(ps.formula, data = analytic.data,
                     distance = 'logit', 
                     method = "nearest", 
                     replace=FALSE,
                     caliper = .2*sd(logitPS), 
                     ratio = 2)
# see matchit function options here
# https://www.rdocumentation.org/packages/MatchIt/versions/1.0-1/topics/matchit
analytic.data$PS <- match.obj$distance
summary(match.obj$distance)
plot(match.obj, type = "jitter")
plot(match.obj, type = "hist")
```

```{r ps2b, cache=TRUE}
tapply(analytic.data$PS, analytic.data$RHC, summary)
# check how many matched
match.obj
summary(match.obj, standardize = TRUE)
# extract matched data
matched.data <- match.data(match.obj)
```

## Step 3

Compare the similarity of baseline characteristics between treated and untreated subjects in a the propensity score-matched sample. In this case, we will compare SMD < 0.1 or not

```{r ps3, cache=TRUE, echo= TRUE,results="hide"}
tab1m <- CreateTableOne(vars = baselinevars,
               data = matched.data, strata = "RHC", includeNA = TRUE, 
               test = FALSE, smd = TRUE)
tab1mk <- kable(print(tab1m, showAllLevels = TRUE, smd = TRUE)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = FALSE)
```

```{r ps3b, cache=TRUE}
tab1mk
```


## Step 4

Estimate the effect of treatment on outcomes using propensity score-matched sample

```{r, ps4, cache=TRUE}
fit3 <- glm(I(Death=="Yes")~RHC,
            family=binomial, data = matched.data)
summ(fit3, exp = TRUE)
publish(fit3)
```

General idea is that if any SMD >= 0.1, go back to the modelling stage, and change the PS model specification. If all else fails, adjust for the variables that are indicating some residual imbalance.

```{r, ps4b, cache=TRUE}
fit4 <- glm(I(Death=="Yes")~RHC+age+sex+Disease.category+
              Blood.pressure+Respiratory.rate+PaCO2,
            family=binomial, data = matched.data)
summ(fit4, exp = TRUE)
publish(fit4)
```
